
\begin{thebibliography}{9}

\bibitem{1}
\url{https://www.hyperledger.org/}

\bibitem{2}
\url{https://gdpr-info.eu/}

\bibitem{3}
\url{https://oag.ca.gov/privacy/ccpa}

\bibitem{4}
\url{https://ethereum.org/en/}

\bibitem{5}
\url{https://bitcoin.org/en/}

\bibitem{6}
 Rahime Belen-Saglam, Enes Altuncu, Yang Lu, Shujun Li. A Systematic Literature Review of the Tension between the GDPR and Public Blockchain Systems. Blockchain: Research and Applications, 2022

 \bibitem{7}
Eugenia Politou, Fran Casino, Efthimios Alepis, Constantinos Patsakis. Blockchain Mutability: Challenges and Proposed Solutions. IEEE Transactions on Emerging Topics in Computing, 2019

\bibitem{8}
Antonio Emerson Barros Tomaz, Jose Caludio Do Nascimento, Abdelhakim Senhaji Hafid, Jose Neuman De Souza. Preserving Privacy in Mobile Health Systems Using Non-Interactive Zero-Knowledge Proof and Blockchain. IEEE Access, 2020

\bibitem{9}
Eugenio Balistria, Francesco Casellatoa, Carlo Giannellib, Cesare Stefanellia. BlockHealth: Blockchain-based secure and peer-to-peer health information sharing with data protection and right to be forgotten. ICT Express, 2021

\bibitem{10}
Aurelie Bayle,  Mirko Koscina, David Manset. When Blockchain Meets the Right to be Forgotten: Technology Versus Law in the Healthcare Industry. IEEE/WIC/ACM International Conference on Web Intelligence (WI), 2018

\bibitem{11}
Xinyu Li, Jing Xu, Lingyuan Yin, Yuan Lu, Qiang Tang, Zhenfeng Zhang. Escaping From Consensus: Instantly Redactable Blockchain Protocols in Permissionless Setting. IEEE Transactions on Dependable and Secure Computing, 2023

\bibitem{12}
Imran Makhdoom, Ian Zhou, Mehran Abolhasan, Justin Lipman, Wei Ni. PrivySharing: A blockchain-based framework for privacy-preserving and secure data sharing in smart cities. Computers \& Security, 2019

\bibitem{13}
Aloni Cohen, Adam Smith, Marika Swanberg, Prashant Nalini Vasudevan. Control, Confidentiality,
and the Right to be Forgotten. arXiv preprint, 2022.

\bibitem{14}
Leslie Lamport. Paxos made simple. ACM SIGACT, 2001

\bibitem{15}
Miguel Castro, Barbara Liskov. Practical byzantine fault tolerance. OSDI, 1999

\bibitem{16}
Diego Ongaro, John Ousterhout. In Search of an Understandable Consensus Algorithm. USENIX ATC, 2014

\bibitem{17}
AES: The Advanced Encryption Standard. Lecture 8 of Computers and Network Security, Purdue University

\bibitem{18}
Public-Key Cryptography and the RSA Algorithm. Lecture 12 of Computers and Network Security, Purdue University

\bibitem{19}
Hashing for Message Authentication. Lecture 15 of Computers and Network Security, Purdue University

\end{thebibliography}
