\documentclass[12pt, a4paper, oneside]{Thesis}
\usepackage{wrapfig}
\usepackage{lscape}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage[T1]{fontenc}
\usepackage{makecell}
\usepackage{changepage}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{array}
\usepackage{float}
\usepackage{placeins}
\usepackage{url}
\usepackage{enumitem}

% remove hyphenation
\tolerance=9999
\emergencystretch=10pt
\hyphenpenalty=10000
\exhyphenpenalty=100

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, positioning, fit, backgrounds, calc, arrows.meta}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}

\setcounter{secnumdepth}{3}
\graphicspath{{Pictures/}}

\usepackage{natbib}

% All links/text BLACK --- no coloured hyperlinks
\hypersetup{urlcolor=black, linkcolor=black, citecolor=black, colorlinks=false}

% ---- Thesis Metadata ----
\thesistitle{Design and Evaluation of Multi-Agent LLM Systems for Interactive Clinical Diagnosis}
\supervisor{Prof. Jiaul Hoque Paik}
\degree{Master of Technology (M.Tech)}
\degreemajor{Department of Artificial Intelligence}
\authors{Dikshant Sharma}
\rollno{21IM3AI29}
\university{Indian Institute of Technology Kharagpur}
\department{Department of Artificial Intelligence}
\unisite{http://www.iitkgp.ac.in}
\depsite{http://www.iitkgp.ac.in}
\placeshrt{Kharagpur}
\placelng{Kharagpur - 721302, India}
\datesub{April, 2026}
\datesig{April, 2026}
\semsub{the academic year 2025--2026}
\keywords{Large Language Models, Clinical Decision-Making, Multi-Agent Systems, LangGraph, Cognitive Bias, Process Evaluation}
\coursecd{Project IV (AI67102)}

\title{\fontsize{13.2pt}{16pt}\selectfont \ttitle}

\begin{document}

\frontmatter
\setstretch{1.6}

\fancyhead{}
\rhead{\thepage}
\lhead{}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\hypersetup{pdftitle={\ttitle}}
\hypersetup{pdfauthor=\authornames}
\hypersetup{pdfkeywords=\keywordnames}

%------------------------------------------------------------------------
% TITLE PAGE
%------------------------------------------------------------------------
\maketitle
\clearpage

%------------------------------------------------------------------------
% DECLARATION
%------------------------------------------------------------------------
\Declaration{}

%------------------------------------------------------------------------
% CERTIFICATE
%------------------------------------------------------------------------
\addtotoc{Certificate}
\certificate{\addtocontents{toc}{}

%------------------------------------------------------------------------
% ABSTRACT
%------------------------------------------------------------------------
\addtotoc{Abstract}
\abstract{\addtocontents{toc}{}

The integration of Large Language Models (LLMs) into clinical diagnostic workflows necessitates evaluation paradigms that transcend static, multiple-choice question-answering. Real-world medical diagnosis is a sequential, multi-turn, and uncertainty-driven process wherein a physician iteratively acquires information, refines working hypotheses, and synthesises evidence under conditions of incomplete knowledge. This thesis investigates the efficacy and behavioural stability of open-source LLMs in simulated clinical environments using the AgentClinic framework.

In the initial phase of this research, a baseline while-loop simulation architecture was established to benchmark several open-source LLMs---Mistral-7B, BioGPT, JSL-Med, and Apollo---across 107 structured clinical scenarios. While open-source models demonstrated baseline clinical competence, with Mistral-7B achieving up to 43.9\% diagnostic accuracy on the AgentClinic-MedQA subset, the unconstrained architecture induced erratic reasoning trajectories and introduced a risk of \textit{Lexical Leakage}---a vulnerability wherein homogeneous multi-agent systems implicitly share ground-truth information through shared model weights.

In the enhanced phase of this research, \textbf{the AgentClinic framework was redesigned} as a multi-agent system employing a LangGraph Directed Cyclic Graph (DCG) for deterministic orchestration. To eliminate Lexical Leakage without exceeding the 95.8\,GB VRAM constraint of an NVIDIA H100 GPU, a heterogeneous multi-engine architecture was implemented, leveraging 4-bit NormalFloat4 (NF4) double quantisation. A latent metacognitive reflection node was introduced, rendering internal diagnostic reasoning observable and quantifiable. Three process-oriented metrics were formally defined: Diagnostic Stability (DS), Test Rationality (TR), and Information Efficiency (IE).

Evaluations on the structured MedQA dataset demonstrate that the structural framework of the DCG and reflection node improved baseline accuracy to \textbf{45.8\%}, indicating that the reasoning gains from forced metacognition more than compensate for the removal of implicit Lexical Leakage. Empirical results further indicate that medically fine-tuned models (JSL-MedMistral) and advanced reasoners (Llama-3.1) improve outcome accuracy, yet remain susceptible to cognitive distortions---recency bias triggering hypothesis instability and confirmation bias inducing premature diagnostic closure. This thesis establishes a trajectory-aware evaluation platform for clinical AI systems.

\clearpage

%------------------------------------------------------------------------
% ACKNOWLEDGEMENTS
%------------------------------------------------------------------------
\setstretch{1.3}
\acknowledgements{\addtocontents{toc}{}

I sincerely thank my project guide, Prof. Jiaul Hoque Paik, for his invaluable insights and constructive feedback, which made this project a successful endeavour. Furthermore, I thank the Department of Artificial Intelligence for granting me the opportunity and resources to pursue this project. I extend my gratitude to the IIT Kharagpur computing infrastructure team for providing access to the NVIDIA H100 NVL GPU cluster that made these experiments possible. I finally extend my deepest gratitude to my family and friends for their unwavering support throughout this journey.

}

%------------------------------------------------------------------------
% TABLE OF CONTENTS / FIGURES / TABLES
%------------------------------------------------------------------------
\clearpage
\pagestyle{fancy}
\lhead{\emph{Contents}}
\tableofcontents


%------------------------------------------------------------------------
% MAIN MATTER
%------------------------------------------------------------------------
\mainmatter
\pagestyle{fancy}

\input{Chapters/1_Introduction}
\input{Chapters/2_Literature_Review}
\input{Chapters/3_Dataset}
\input{Chapters/4_Initial_Framework}
\input{Chapters/5_Enhanced_Methodology}
\input{Chapters/6_Results}
\input{Chapters/7_Discussion}
\input{Chapters/8_Conclusion}

\appendix
\input{Chapters/Appendix_A}

%------------------------------------------------------------------------
% BIBLIOGRAPHY
%------------------------------------------------------------------------
\label{References}
\lhead{\emph{Bibliography}}
\nocite{*}
\bibliographystyle{unsrt}
\bibliography{Bibliography}

\end{document}
