\documentclass[12pt, a4paper, oneside]{Thesis}
\usepackage{wrapfig}
\usepackage{lscape}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage[T1]{fontenc}
\usepackage{makecell}
\usepackage{changepage}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{array}
\usepackage{float}
\usepackage{placeins}
\usepackage{url}
\usepackage{enumitem}

% remove hyphenation
\tolerance=9999
\emergencystretch=10pt
\hyphenpenalty=10000
\exhyphenpenalty=100

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, positioning, fit, backgrounds, calc, arrows.meta}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}

\setcounter{secnumdepth}{3}
\graphicspath{{Pictures/}}

\usepackage{natbib}

% All links/text BLACK — no coloured hyperlinks
\hypersetup{urlcolor=black, linkcolor=black, citecolor=black, colorlinks=false}

% ---- Thesis Metadata ----
\thesistitle{Design and Evaluation of Multi-Agent LLM Systems for Interactive Clinical Diagnosis}
\supervisor{Prof. Jiaul Hoque Paik}
\degree{Masters of Technology}
\degreemajor{Department of Artificial Intelligence}
\authors{Dikshant Sharma}
\rollno{21IM3AI29}
\university{Indian Institute of Technology Kharagpur}
\department{Department of Artificial Intelligence}
\unisite{http://www.iitkgp.ac.in}
\depsite{http://www.iitkgp.ac.in}
\placeshrt{Kharagpur}
\placelng{Kharagpur - 721302, India}
\datesub{April, 2026}
\datesig{April, 2026}
\semsub{}
\keywords{Large Language Models, Clinical Decision-Making, Multi-Agent Systems, LangGraph, Cognitive Bias, Process Evaluation}
\coursecd{Project-III (AI67101)}

\title{\fontsize{13.2pt}{16pt}\selectfont \ttitle}

\begin{document}

\frontmatter
\setstretch{1.6}

\fancyhead{}
\rhead{\thepage}
\lhead{}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\hypersetup{pdftitle={\ttitle}}
\hypersetup{pdfauthor=\authornames}
\hypersetup{pdfkeywords=\keywordnames}

%------------------------------------------------------------------------
% TITLE PAGE
%------------------------------------------------------------------------
\maketitle
\clearpage

%------------------------------------------------------------------------
% DECLARATION
%------------------------------------------------------------------------
\Declaration
\clearpage

%------------------------------------------------------------------------
% CERTIFICATE
%------------------------------------------------------------------------
\addtotoc{Certificate}
\certificate{\addtocontents{toc}{}

\clearpage

%------------------------------------------------------------------------
% ABSTRACT
%------------------------------------------------------------------------
\addtotoc{Abstract}
\abstract{\addtocontents{toc}{}

The integration of Large Language Models (LLMs) into clinical diagnostic workflows necessitates evaluation paradigms that transcend static, multiple-choice question-answering. Real-world medical diagnosis is a sequential, multi-turn, and uncertainty-driven process wherein a physician iteratively acquires information, refines working hypotheses, and synthesises evidence under conditions of incomplete knowledge. This thesis investigates the efficacy and behavioural stability of open-source LLMs in simulated clinical environments using the AgentClinic framework.

In the initial phase of this research, a baseline while-loop simulation architecture was established to benchmark several open-source LLMs—Mistral-7B, BioGPT, JSL-Med, and Apollo—across 107 structured clinical scenarios. Whilst open-source models demonstrated baseline clinical competence, with Mistral-7B achieving up to 44\% diagnostic accuracy on the AgentClinic-MedQA subset, the unconstrained architecture induced erratic reasoning trajectories and posed a risk of \textit{Lexical Leakage}—a phenomenon wherein homogeneous multi-agent systems may subconsciously share latent ground-truth representations through overlapping embedding manifolds.

In the enhanced phase of this research, \textbf{AgentClinic v2} was engineered: an advanced multi-agent framework employing a LangGraph Directed Cyclic Graph (DCG) for deterministic orchestration. To eliminate Lexical Leakage without exceeding the 95.8\,GB VRAM constraint of an NVIDIA H100 GPU, a heterogeneous multi-engine architecture was implemented, leveraging 4-bit NormalFloat4 (NF4) double quantization. Crucially, a latent metacognitive reflection node was introduced, rendering internal diagnostic reasoning observable and quantifiable. Three novel process-oriented metrics were formally defined: Diagnostic Stability (DS), Test Rationality (TR), and Information Efficiency (IE).

Evaluations on the structured MedQA dataset reveal a profound finding: the structural scaffolding of the DCG and reflection node improved the baseline accuracy to \textbf{45.6\%}, demonstrating that forced metacognition entirely eclipses the performance penalty of removing implicit Lexical Leakage. Empirical validations further indicate that medically fine-tuned models (JSL-MedMistral) and advanced reasoners (Llama-3.1) improve outcome accuracy, yet remain susceptible to cognitive distortions—recency bias triggering catastrophic hypothesis thrashing and confirmation bias inducing premature diagnostic closure. This thesis establishes a comprehensive, trajectory-aware platform for the robust evaluation of clinical AI systems.

\clearpage

%------------------------------------------------------------------------
% ACKNOWLEDGEMENTS
%------------------------------------------------------------------------
\setstretch{1.3}
\acknowledgements{\addtocontents{toc}{}

I sincerely thank my project guide, Prof. Jiaul Hoque Paik, for his invaluable insights and constructive feedback, which made this project a successful endeavour. Furthermore, I thank the Department of Artificial Intelligence for granting me the opportunity and resources to pursue this project. I extend my gratitude to the IIT Kharagpur computing infrastructure team for providing access to the NVIDIA H100 NVL GPU cluster that made these experiments possible. I finally extend my deepest gratitude to my family and friends for their unwavering support throughout this journey.

}

%------------------------------------------------------------------------
% TABLE OF CONTENTS / FIGURES / TABLES
%------------------------------------------------------------------------
\clearpage
\pagestyle{fancy}
\lhead{\emph{Contents}}
\tableofcontents

\clearpage
\lhead{\emph{List of Figures}}
\listoffigures

\clearpage
\lhead{\emph{List of Tables}}
\listoftables

%------------------------------------------------------------------------
% MAIN MATTER
%------------------------------------------------------------------------
\mainmatter
\pagestyle{fancy}

\input{Chapters/1_Introduction}
\input{Chapters/2_Literature_Review}
\input{Chapters/3_Dataset}
\input{Chapters/4_Initial_Framework}
\input{Chapters/5_Enhanced_Methodology}
\input{Chapters/6_Results}
\input{Chapters/7_Discussion}
\input{Chapters/8_Conclusion}

%------------------------------------------------------------------------
% BIBLIOGRAPHY
%------------------------------------------------------------------------
\label{References}
\lhead{\emph{Bibliography}}
\nocite{*}
\bibliographystyle{unsrt}
\bibliography{Bibliography}

\end{document}
